---
title: "tunnel_fb_analyses"
author: "Lucrezia Lonardo"
date: "2025-06-28"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(lme4)
library(ggsignif)
library(summarytools)
library(car)
source("functions/diagnostic_fcns.r")
source("functions/glmm_stability.r")
source("functions/boot_glmm.r")
#load("~/GitHub/tunnel_fb/workspace/Tunnel_FB_workspace.RData")
```

# Import data

```{r cars}
#import choice data
xdata <- read.csv("data/Tunnel_FB_tested_dogs.csv", header = TRUE) %>% 
   mutate(first_touch_letter = as.factor(first_touch_letter),
         condition = as.factor(condition),
         age_months = as.numeric(age_months)) %>% 
  filter(first_touch_letter %in% c("A", "B")) %>%   # Keep rows where first_touch_letter is "A" or "B" in case of NAs
  filter(CDL_name!="Vanilla2") %>% #should be in the excluded tab, did not meet criteria in fam. phase 1
  droplevels()

which(is.na(xdata$first_touch_letter)) #double check for NAs

#import latency data
ldata<-read.csv("data/Tunnel_FB_latency.csv", header = TRUE) %>% 
  select(where(~ any(!is.na(.)))) %>%  #keeps only columns where at least one value is not NA (gets rid of columns containing only NAs) %>% 
  select(-c("X", "Scoring", "Start_Frame", "Stop_Frame", "Value"))
```

## Obtain durations from latency file

```{r}
ldata<-ldata %>% 
  mutate(duration=Stop-Start) %>% 
  select(-c("Start", "Stop")) %>% 
  pivot_wider(values_from = duration, names_from = Behaviour)
```

## Join choice and latency data
```{r}
all.data<-xdata %>% 
  rename(Subject=CDL_name) %>% 
 left_join(ldata, by = "Subject") #left join because currently not all dogs have been coded
```


# Plots

###Barplot all breeds
```{r pressure, echo=FALSE}
plot.data <- xdata %>%
  mutate(choice = ifelse(first_touch_letter == "A", 1, 0)) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "false belief" = "false_belief",
      "true belief" = "true_belief"
    ),
    "false belief",
    "true belief"
  )) %>% #specify order of factor levels for factor_relevel
  group_by(condition) %>% 
  summarise(
    mean.choiceA = mean(choice),
    sd.choiceA = sd(choice),
    n.choseA = sum(choice),
    median.choiceA = median(choice),
    n.tested.per.cond = length(choice),
    se.choiceA = sd(choice) / sqrt(length(choice))
  )

##barplot of choices by experimental group

p.choice <- ggplot(data = plot.data, aes(x = condition, y = mean.choiceA)) +
  geom_bar(stat = "identity", alpha = 0.5) +
  geom_errorbar(aes(ymin = mean.choiceA - se.choiceA, ymax = mean.choiceA + #error bars represent standard error
                      se.choiceA),
                width = .2) +
  #geom_signif(comparisons=list(c("False Belief", #"Knowledge")), annotations="*",
  #             y_position = 0.75, tip_length = 0.04, #vjust=0.4) +
  #ylim(0.2,0.6)+
  theme_classic() +
  labs(x = "Experimental group", y = "Proportion of bucket A choices")+
  ggtitle("All breeds (N=63)")

p.choice
```
```{r}
#save the choices plot

ggsave(
  "./plots/bucket_A_choices_all_breeds.png",
  plot = p.choice,
  device = NULL,
  scale = 0.8,
  width = 7,
  height = 6,
  #units = c("in", "cm", "mm"),
  dpi = 320,
  limitsize = TRUE
)
```


```{r}
#add column with independent vs cooperative breeds distinction 
xdata<-xdata %>% 
  mutate(breed_type = ifelse(FCI_group %in% c("1", "2", "7", "8"), "cooperative", "independent"))

table(xdata$breed_type)

table(xdata$first_touch_letter, xdata$breed_type, xdata$condition)

# #add column with experience in dogs sport as binary (yes/no) variable
# xdata<-xdata %>% 
#   mutate(training = ifelse(experience_in_dog_sport=="no", "no", "yes"))
# 
# table(xdata$training)
# 
# table(xdata$FCI_group[xdata$training=="no"])
# 
# table(xdata$first_touch_letter, xdata$training, xdata$condition)

#experimenter ID
table(xdata$communicator, xdata$first_touch_letter, xdata$condition)
```
### Plot breed performance
```{r}
cooperative_data<-xdata %>% 
  filter(breed_type!="independent")
independent_data<-xdata %>% 
  filter(breed_type!="cooperative")

#mosaic plot data for cooperative
mosaic.cooperative.plot.data  <- cooperative_data %>%
  mutate(first_touch_letter = fct_relevel(first_touch_letter, "B", "A")) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "False belief" = "false_belief",
      "True Belief" = "true_belief"
    ),
    "False belief",
    "True Belief"
  )) #specify order of factor levels for factor_relevel

#mosaic plot data for independent
mosaic.independent.plot.data  <- independent_data %>%
  mutate(first_touch_letter = fct_relevel(first_touch_letter, "B", "A")) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "False belief" = "false_belief",
      "True Belief" = "true_belief"
    ),
    "False belief",
    "True Belief"
  )) #specify order of factor levels for factor_relevel


#create and save the plot
png("./plots/breed_type_mosaic_plot.png", width = 1600, height = 800) 
par(mfrow=c(1, 2), mar=c(6, 7.8, 6, 5), mgp=c(4.5, 1.5, 0), tcl=-0.15, las=1,
    cex.axis = 2.6,   # Increase axis tick labels
    cex.lab = 2.9,    # Increase axis labels
    cex.main = 3.5)     # Increase plot titles

plot(mosaic.cooperative.plot.data$condition,mosaic.cooperative.plot.data$first_touch_letter,
  xlab = "Condition",
  ylab = "Bucket chosen",
  main = "Cooperative breeds" )

plot(mosaic.independent.plot.data$condition,mosaic.independent.plot.data$first_touch_letter,
  xlab = "Condition",
  ylab = "Bucket chosen",
  main = "Independent breeds" )
dev.off()

```
### Plot latencies
```{r}
hist(ldata$man_check_duration)
hist(ldata$latency)


plot.latency.data <- all.data %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "false belief" = "false_belief",
      "true belief" = "true_belief"
    ),
    "false belief",
    "true belief"
  )) %>% #specify order of factor levels for factor_relevel
group_by(condition) %>%
  summarise(
    mean.latency = mean(latency, na.rm = TRUE),
    sd.latency = sd(latency, na.rm = TRUE),
    mean.man.check.dur = mean(man_check_duration, na.rm = TRUE),
    se.latency = sd(latency, na.rm = TRUE) / sqrt(sum(!is.na(latency)))
  )


##barplot of choices by experimental group

p.latency <- ggplot(data = plot.latency.data, aes(x = condition, y = mean.latency)) +
  geom_bar(stat = "identity", alpha = 0.5) +
  geom_errorbar(aes(ymin = mean.latency - se.latency, ymax = mean.latency + se.latency), #error bars represent standard error
                width = .2) +
  geom_signif(comparisons=list(c("false belief", "true belief")), annotations="*",
               y_position = 26, tip_length = 0.02, vjust=0.4) +
  #ylim(0.2,0.6)+
  theme_classic() +
  labs(x = "Experimental group", y = "Latency to approach the tunnel (s)")+
  ggtitle("All breeds (N=63)")

p.latency
```
```{r}
#save the latency plot

ggsave(
  "./plots/latency_to_approach_tunnel.png",
  plot = p.latency,
  device = NULL,
  scale = 0.8,
  width = 7,
  height = 6,
  #units = c("in", "cm", "mm"),
  dpi = 320,
  limitsize = TRUE
)
```

# Check how often dogs switched choice after tunnel
```{r}
all.data <- all.data %>%
  mutate(change_of_choice = case_when(
    first_choice_letter == "A" & first_touch_letter == "B" ~ "A-B",
    first_choice_letter == "B" & first_touch_letter == "A" ~ "B-A",
    first_choice_letter == first_touch_letter ~ "no change",
    TRUE ~ NA_character_  # handles unexpected combinations
  ))
```


#Analyse dogs' choices

##Prepare data
```{r}
levels(as.factor(all.data$breed)) #some levels need recoding
all.data<-all.data %>% 
    mutate(breed_recoded = case_when(
    breed == "wolfspitz" 
  #  | breed == "spitz" 
    ~ "german_spitz", #LL: I am assuming spitz was also German spitz
    TRUE ~ as.character(breed) # Preserve original values as character
    ),
  breed_recoded = as.factor(breed_recoded))

levels(as.factor(all.data$breed_recoded)) 

 model.data<-all.data %>% 
  mutate(breed_type = ifelse(FCI_group %in% c("1", "2", "7", "8"), "cooperative", "independent")) %>%  #add breed type
  #making sure factors are treated as such
  mutate(Subject=as.factor(Subject)) %>% 
  mutate(breed=as.factor(breed)) %>% 
  mutate(breed_type=as.factor(breed_type)) %>% 
  mutate(sex=as.factor(sex)) %>% 
  mutate(castrated=as.factor(castrated)) %>% 
  mutate(first_baited_bucket=as.factor(first_baited_bucket)) %>% 
  # select(-first_bucket_A) %>% 
  mutate(FCI_group=as.factor(FCI_group)) %>% 
  #mutate(training=as.factor(training)) %>%
  mutate(hider=as.factor(hider)) %>%
  mutate(communicator=as.factor(communicator)) %>%
  #mutate(experimenter_dyad=as.factor(experimenter_dyad)) %>%
  mutate(first_touch_letter=as.factor(first_touch_letter)) %>% 
  #centering variables
  mutate(z.age = as.numeric(scale(age_months, scale = T, center = T)), #age scaled and centered (=z-transformed, we subtracted the mean and divided by SD)
         sex.c = as.numeric(scale(as.numeric(as.factor(sex)), scale=F, center=T)), #factors manually dummy coded and centered (i.e., we subtracted the mean value from each observed value)
         condition.c = as.numeric(scale(as.numeric(as.factor(condition)), scale=F, center=T)),
         breed_type.c = as.numeric(scale(as.numeric(as.factor(breed_type)), scale=F, center=T)),
         first_baited_bucket.c = as.numeric(scale(as.numeric(as.factor(first_baited_bucket)), scale=F, center=T)),
         #training.c = as.numeric(scale(as.numeric(as.factor(training)), scale=F, center=T))
         )



view(dfSummary(model.data))

summary(model.data$age_months)

#checking familiarisation performance is as expected (max 4 trials per phase)
which(model.data$nr_fam_phase1_trials>4) 

#check new levels of breed 
unique(levels(model.data$breed_recoded)) #17 different breeds

#checking if the data preparations worked as expected
str(model.data)
```
### GLMM 01: choice - model with breed as random intercept
```{r}
xx.fe.re=fe.re.tab(fe.model="first_touch_letter ~ condition*breed_type + age_months + sex + first_baited_bucket",
                   re="(1|breed_recoded)", data=model.data)
xx.fe.re$summary
```

``` {r mixed modeling, error=TRUE}
#full model with all rs within breed

full.choice.int <- glmer(first_touch_letter ~ condition*breed_type + z.age + sex + first_baited_bucket+
             (1+ sex.c + condition.c + first_baited_bucket.c + z.age || breed_recoded),
             data = model.data, family = binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
)

#had to exclude correlation parameters between rss and icpt because they were unidentifiable
```

#### model output
  + Coefficients
```{r}
round(summary(full.choice.int)$coefficients, 3)
```
  + Individual predictors: likelihood ratio tests  
Drop1: P values for the individual effects were based on likelihood ratio tests comparing the full with respective reduced models (Barr et al., 2013; R function drop1 with argument 'test' set to "Chisq"). 

```{r}
drop1.full=drop1(full.choice.int, test="Chisq")
round(drop1.full,3)
#interaction not significant
```
####the interaction condition*breed_type is not significant, so we remove it but keep both main effects
``` {r mixed modeling, error=TRUE}
#full model with all rs within breed

full.choice.breed <- glmer(first_touch_letter ~ condition + breed_type + z.age + sex + first_baited_bucket+
             (1+ sex.c + condition.c + first_baited_bucket.c + z.age || breed_recoded),
             data = model.data, family = binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
)

#had to exclude correlation parameters between random slopes and intercept because they were unidentifiable
```
#### model output
  + Coefficients
```{r}
round(summary(full.choice.breed)$coefficients, 3)
```
  + Individual predictors: likelihood ratio tests  
Drop1: P values for the individual effects were based on likelihood ratio tests comparing the full with respective reduced models (Barr et al., 2013; R function drop1 with argument 'test' set to "Chisq"). 

```{r}
drop1.full=drop1(full.choice.breed, test="Chisq")
round(drop1.full,3)
#the interaction condition*breed_type is not significant, so we remove it but keep both main effects
```
#### confidence intervals

```{r}
boot.res.breed=boot.glmm.pred(model.res=full.choice.breed, excl.warnings=T,
	nboots=1000, para=T)
res.breed<-round(boot.res.breed$ci.estimates, 2)
res.breed
```

#### model stability

```{r eval=FALSE, include=FALSE}

# One breed at a time excluded to assess the impact of possible outliers 
m.stab.breed <- glmm.model.stab(model.res = full.choice.breed, use = c("breed_recoded"))

table(m.stab.breed$detailed$lme4.warnings) #several models failed to converge

xx.breed <- as.data.frame(round(m.stab.breed$summary[, -1], 2))
#table with model stability
xx.breed

#plot of stability (remember to execute all three lines at the same time)
png("plots/bin_glmm01_breed_type_stability.png")
m.stab.plot(round(m.stab.breed$summary[, -1], 3))
dev.off()
```

#### Variance Inflation Factors
```{r}
vif(full.choice.breed) #collinearity is no problem
```

#### Distribution of BLUPs
```{r}
ranef.diagn.plot(full.choice.breed)
```


#Analyse latency
```{r}
t.test(all.data$latency[all.data$condition=="false_belief"], all.data$latency[all.data$condition=="true_belief"], paired=FALSE)

#check mean man.check.dur
t.test(all.data$man_check_duration[all.data$condition=="false_belief"], all.data$man_check_duration[all.data$condition=="true_belief"], paired=FALSE) #careful! These are almost different, with TB being systematically longer than FB!!
```


